---
title: "bentley tidyverse submission"
author: "AJ Bentley"
date: "November 27, 2018"
output: html_document
---

```{r setup, include=FALSE, echo=TRUE, messages = FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

### Goal

Using the fivethirtyeigth masculinity survey I'm going to use dplyr's select and recode_factor features.

#### Load and examine data

```{r read in data}

raw <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/masculinity-survey/raw-responses.csv", header = TRUE)

raw[1:10, 2:7]

```

#### Getting list of levels for Q1 (q0001)

```{r check unique}

levels(raw$q0001)

```

#### Creating a working df that excludes the first column of raw data (which is just indices)

```{r create working df}

# load required tidyverse library
library(dplyr)

refined <- select(raw, 2:98)

```

#### Recode factors in Q1 responses to numeric / ordered


```{r change q1 to ordered}

# create list of what we want the changes to be
level_key <- list(`No answer` = 1, `Not at all masculine` = 2, `Not very masculine` = 3, `Somewhat masculine` = 4, `Very masculine` = 5)


# apply the recode_factor code
refined$q0001 <- recode_factor(refined$q0001, !!!level_key)

# verify changes were made in column
levels(refined$q0001)

# verify changes were made in df
raw[1:10, 2:7]

```

```{r}
#section added by Chad Bailey
#taking the suggested edit of adding a new dummy column with the numeric value in addition to rather than in place of the original value
refined2 <- raw %>%
              mutate(q0001_coded = ifelse (q0001 == 'No answer', 1,
                                   ifelse (q0001 == 'Not at all masculine', 2,
                                   ifelse (q0001 == 'Not very masculine', 3,
                                   ifelse (q0001 == 'Somewhat masculine', 4,
                                   ifelse (q0001 == 'Very masculine', 5, -99)))))) %>%
              select(StartDate, EndDate, q0001, q0001_coded)

head(refined2)

```


### Suggested Improvements

Recode is far more complex than it needs be. There's no reason to have the list and recode functions be separate.


### Potential steps for other students

Obviously there are plenty of other opportunities to recode. 

Other options would include creating dummy variables for columns that are factors but not ordered.



